---
title: "CensusNHS_DA_income"
author: "Martin Monkman"
date: "March 20, 2017"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Socio-demographic statistics by Dissemination Area

see [2011 National Household Survey: Data tables](http://www12.statcan.gc.ca/nhs-enm/2011/dp-pd/dt-td/Lp-eng.cfm?LANG=E&APATH=3&DETAIL=0&DIM=0&FL=A&FREE=0&GC=0&GID=0&GK=0&GRP=0&PID=0&PRID=0&PTYPE=105277&S=0&SHOWALL=Yes&StartRow=1&SUB=0&Temporal=2013&THEME=98&VID=0&VNAMEE=&VNAMEF=)


For variable names, see file NHS2011_BC.csv (in same folder). The variables listed are shown as the row names in the main file.

```{r}
library(dplyr)
library(readr)
library(stringr)
library(tidyr)


library(ggplot2)


```



read source file and filter so only Dissemination Area stats remain

```{r  message = FALSE}

# NHS2011data <- read_csv("E:/github/housing-data-visualization-project/data/R3_SCP_Can_PR_CD_CSD_DA_AllGNRs_withSUPP_BC.csv")


#NHS2011data <- read_csv("E:/github/housing-data-visualization-project/data/R3_SCP_Can_PR_CD_CSD_DA_AllGNRs_withSUPP_BC.csv",
#                        col_types = cols(.default = "d",
#                                         Geography = "c"))

NHS2011data <- read_csv("C:/@data/GitHub/housing-data-visualization-project/data/R3_SCP_Can_PR_CD_CSD_DA_AllGNRs_withSUPP_BC.csv",
                        col_types = cols(.default = "d",
                                         Geography = "c"))

NHS2011data_DA <- NHS2011data %>%
  filter(str_detect(Geography, '59'))



```


---

#### Housing & related

ranges (columns) of interest:

* mobility status  1718:1771

* housing suitability  2536:2538
* shelter costs  2539:2542/2549

---

#### Labour force

ranges (columns) of interest:

* 1988:2011

* class of worker  2012:2026
* NOC  


---

#### Income and related

ranges (columns) of interest:

* 2231:2383

* median 2247
* average 2248

* employment income 2384:2407
* distribution 2408:2446
* LICO 2447-2491
  - LICO count 2452
  - LICO percentage 2457

* deciles  2514:2520
* total household income (distribution)  2555:2587
  - one person households 2588:2592
  - two+ person households 2593:2597



```{r}



# decile
# 2408: Total population by decile of adjusted after-tax family income

pop_decile <- NHS2011data_DA %>%
  select(1:2,2408:2420)

head(pop_decile)

# LICO

pop_inlowincome <- NHS2011data_DA %>%
  select(1:2,2447:2491)

head(pop_inlowincome)

```


### summary stats


```{r}

# http://stackoverflow.com/questions/22842232/dplyr-select-column-names-containing-white-space
# BACK TICKS!!!


# combo of average, median, and prevelance of low income

pop_income_summary <- NHS2011data_DA %>%
  select(1:2,2514,2247,2248,2457) %>%
  rename(Income_Avg = `Average income $`,
         Income_Med = `Median income $`,
         Prevalance_Low_Income = `Prevalence of low income in 2010 based on after-tax low-income measure %`)


write_csv(pop_income_summary, "pop_income_summary.csv")

```


#### some charts of the above summary

```{r}

# omit na from ggplot2
# http://stackoverflow.com/questions/17216358/eliminating-nas-from-a-ggplot
#ggplot(data=subset(iris, !is.na(Sepal.Length)), aes(x=Sepal.Length)) + 
#geom_bar(stat="bin")


#chart_data <- pop_income_summary %>%
#  select(IncMed) %>%
#  filter(IncMed != is.na())

ggplot(pop_income_summary, aes(x = Income_Med)) +
  geom_histogram(bins = 25)


ggplot(pop_income_summary, aes(x = Income_Avg)) +
  geom_histogram(bins = 25)

ggplot(pop_income_summary, aes(x = Prevalance_Low_Income)) +
  geom_histogram(bins = 25)


ggplot(pop_income_summary, aes(x = Prevalance_Low_Income, y = Income_Med)) +
  geom_point()

```


#### Ranking and Index

**Rank:** by prevalence of low income (high prevalence = high score) and median income (low income = high score.)

**Index:** sum of prevalence of low income rank and median income rank.
A low score means a combination of both _high_ prevalence and _low_ median income.



```{r}

pop_income_summary_02 <- pop_income_summary %>%
  na.omit() %>%
  mutate(Income_Med_rank = rank(desc(Income_Med)),
         Prevalance_Low_Income_rank = rank(Prevalance_Low_Income),
         index = (Income_Med_rank + Prevalance_Low_Income_rank))

ggplot(pop_income_summary_02, aes(x = Prevalance_Low_Income_rank, y = Income_Med_rank)) +
  geom_point()





# Define target areas: 
# Where prevalence of low income is greater than 50%, or where the median income is below 20000.


pop_income_summary_02 <- pop_income_summary_02 %>%
  mutate(target = ifelse(test = (Prevalance_Low_Income > 40 | index > 12000),
                         yes = 1, 
                         no = 0))
  
# create stand alone DA number in pop_income_summary_02

pop_income_summary_02$DAUID <- as.integer(str_sub(pop_income_summary_02$Geography, 1, 8))


```


#### Plot with colour to show target areas

```{r}

ggplot(pop_income_summary_02, aes(x = Prevalance_Low_Income_rank, y = Income_Med_rank,
                                  colour = target)) +
  geom_point()



ggplot(pop_income_summary_02, aes(x = Prevalance_Low_Income, y = Income_Med,
                                  colour = target)) +
  geom_point()



```


#### THIS DOESN'T WORK PROPERLY!

Fixed the double counting but now it's falling short of the poper total. 


```{r}

pop_income_summary_02 %>%
  group_by(target) %>%
  summarise(count = n(),
            pop = sum(`Total population in private households by citizenship`),
            households = sum(`Total number of private households by tenure`))

```



### Add geography codes

```{r}

# read geog file -- with 3 digit FSA

#geog_data <- read_csv("E:/github/housing-data-visualization-project/data/geography/DB2011_all_translations_adjusted_WithCdCmaAddups.csv")
geog_data <- read_csv("C:/@data/GitHub/housing-data-visualization-project/data/geography/DB2011_all_translations_adjusted_WithCdCmaAddups.csv")


geog_data_DA <- geog_data %>%
  distinct(DAUID, FSA, CDUID, CSDUID, CSDNAME, CSDTYPE)

ls.str(geog_data_DA)


pop_income_summary_03 <- pop_income_summary_02 %>% 
  left_join(geog_data_DA, by = "DAUID") %>%
  arrange(desc(index))


write_csv(pop_income_summary_03, "pop_income_summary_03.csv")


```




```{r}

# read geog file -- with 6 digit postal code

GCS_201702 <- read_csv("C:/@data/GitHub/housing-data-visualization-project/data/geography/GCS_201702.csv")

# test on FERNIE 
#FERNIE <- GCS_201702 %>%
#  filter(COMM_NAME == "FERNIE")
#
#FERNIE <- FERNIE %>%
#  mutate(DAUID = as.integer(paste("59", CD_2011, DA_2011, sep = ''))) 
#
#ls.str(FERNIE)

geog_data_2 <- GCS_201702 %>%
  mutate(DAUID = as.integer(paste("59", CD_2011, DA_2011, sep = ''))) 
  
geog_data_2 <- geog_data_2 %>%
  distinct(DAUID, POSTALCODE, CD_2011, CSD_2011, COMM_NAME)

ls.str(geog_data_2)


pop_income_summary_04 <- pop_income_summary_02 %>% 
  left_join(geog_data_2, by = "DAUID") %>%
  arrange(desc(index))

ls.str(pop_income_summary_04)

write_csv(pop_income_summary_04, "pop_income_summary_04.csv")


```

