---
title: "CensusNHS_DA_income"
author: "Martin Monkman"
date: "March 20, 2017"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Socio-demographic statistics by Dissemination Area

see [2011 National Household Survey: Data tables](http://www12.statcan.gc.ca/nhs-enm/2011/dp-pd/dt-td/Lp-eng.cfm?LANG=E&APATH=3&DETAIL=0&DIM=0&FL=A&FREE=0&GC=0&GID=0&GK=0&GRP=0&PID=0&PRID=0&PTYPE=105277&S=0&SHOWALL=Yes&StartRow=1&SUB=0&Temporal=2013&THEME=98&VID=0&VNAMEE=&VNAMEF=)


For variable names, see file NHS2011_BC.csv (in same folder). The variables listed are shown as the row names in the main file.

```{r}
library(dplyr)
library(readr)
library(stringr)
library(tidyr)


library(ggplot2)


```



read file

```{r  message = FALSE}

# NHS2011data <- read_csv("E:/github/housing-data-visualization-project/data/R3_SCP_Can_PR_CD_CSD_DA_AllGNRs_withSUPP_BC.csv")


NHS2011data <- read_csv("E:/github/housing-data-visualization-project/data/R3_SCP_Can_PR_CD_CSD_DA_AllGNRs_withSUPP_BC.csv",
                        col_types = cols(.default = "d",
                                         Geography = "c"))


```

#### Filter so only Dissemination Area stats remain

```{r}

NHS2011data_DA <- NHS2011data %>%
  filter(str_detect(Geography, '59'))


```



#### Income and related

```{r}



# decile
# 2408: Total population by decile of adjusted after-tax family income

pop_decile <- NHS2011data_DA %>%
  select(1:2,2408:2420)

head(pop_decile)

# LICO

pop_inlowincome <- NHS2011data_DA %>%
  select(1:2,2447:2491)

head(pop_inlowincome)

```


### summary stats


```{r}

# http://stackoverflow.com/questions/22842232/dplyr-select-column-names-containing-white-space
# BACK TICKS!!!


# combo

pop_income_summary <- NHS2011data_DA %>%
  select(1:2,2514,2247,2248,2457) %>%
  rename(Income_Avg = `Average income $`,
         Income_Med = `Median income $`,
         Prevalance_Low_Income = `Prevalence of low income in 2010 based on after-tax low-income measure %`)


write_csv(pop_income_summary, "pop_income_summary.csv")

```


#### some charts of the above summary

```{r}

# omit na from ggplot2
# http://stackoverflow.com/questions/17216358/eliminating-nas-from-a-ggplot
#ggplot(data=subset(iris, !is.na(Sepal.Length)), aes(x=Sepal.Length)) + 
#geom_bar(stat="bin")


#chart_data <- pop_income_summary %>%
#  select(IncMed) %>%
#  filter(IncMed != is.na())

ggplot(pop_income_summary, aes(x = Income_Med)) +
  geom_histogram(bins = 25)


ggplot(pop_income_summary, aes(x = Income_Avg)) +
  geom_histogram(bins = 25)

ggplot(pop_income_summary, aes(x = Prevalance_Low_Income)) +
  geom_histogram(bins = 25)


ggplot(pop_income_summary, aes(x = Prevalance_Low_Income, y = Income_Med)) +
  geom_point()

```


#### Ranking and Index

**Rank:** by prevalence of low income (high prevalence = high score) and median income (low income = high score.)

**Index:** sum of prevalence of low income rank and median income rank.
A low score means a combination of both _high_ prevalence and _low_ median income.



```{r}

pop_income_summary_02 <- pop_income_summary %>%
  na.omit() %>%
  mutate(Income_Med_rank = rank(desc(Income_Med)),
         Prevalance_Low_Income_rank = rank(Prevalance_Low_Income),
         index = (Income_Med_rank + Prevalance_Low_Income_rank))

ggplot(pop_income_summary_02, aes(x = Prevalance_Low_Income_rank, y = Income_Med_rank)) +
  geom_point()



```


#### Define target areas

Where prevalence of low income is greater than 50%, or where the median income is below 20000.

```{r}

pop_income_summary_02 <- pop_income_summary_02 %>%
  mutate(target = ifelse(test = (Prevalance_Low_Income > 40 | index > 12000),
                         yes = 1, 
                         no = 0))
  


```


#### Plot with colour to show target areas

```{r}

ggplot(pop_income_summary_02, aes(x = Prevalance_Low_Income_rank, y = Income_Med_rank,
                                  colour = target)) +
  geom_point()



ggplot(pop_income_summary_02, aes(x = Prevalance_Low_Income, y = Income_Med,
                                  colour = target)) +
  geom_point()



```


#### THIS DOESN'T WORK PROPERLY!

Fixed the double counting but now it's falling short of the poper total. 


```{r}

pop_income_summary_02 %>%
  group_by(target) %>%
  summarise(count = n(),
            pop = sum(`Total population in private households by citizenship`),
            households = sum(`Total number of private households by tenure`))

```

